---
title: "CAMP-R Notebook"
output: html_notebook
---

```{r}
# for data manipulation and visualization
library(tidyverse)

# to help set the working directory and make this script reproducible across different machines
library(here)

# for the strobe diagram
library(DiagrammeR)
library(DiagrammeRsvg) 
library(glue)
library(rsvg)

library(tableone) 

```
## check the working directory- this should be the filepath to your project folder

```{r}
here::here()
```
Load in the data set
```{r}
ped_health <- read_csv(here("ped_health.csv"))
view(ped_health)
```
Check distribution Selecteed child age in years
```{r}
summary(ped_health$SC_AGE_YEARS)
table(ped_health$SC_AGE_YEARS)
```



Check distribution of those with adhd 
```{r}
summary(ped_health$K2Q31B)
table(ped_health$K2Q31B)

### 1- yes, 2- no, 95- don't know/refused, 99- missing
```
Recode this variable to be a factor....
```{r}
ped_health <- ped_health %>%
  mutate(diagnosed_adhd = case_when(
    K2Q31A == 1 ~ "Yes",
    K2Q31A == 2 ~ "No",
    K2Q31A == 95 ~ "Don't know/refused",
    K2Q31A == 99 ~ "Missing"
  )) %>%
  mutate(diagnosed_adhd = factor(adhd, levels = c("Yes", "No", "Don't know/refused", "Missing")))
```
Visualize this distribution
```{r}
ggplot(ped_health, aes(x = diagnosed_adhd)) +
  geom_bar() +
  labs(title = "Distribution of ADHD Diagnosis", x = "ADHD Diagnosis", y = "Count") +
  theme_minimal() +
  scale_x_discrete(drop = FALSE) # to ensure all factor levels are shown
```
Now let's look at the distribution of those born with cardiac illness.
```{r}
summary(ped_health$HeartBorn_2223)
table(ped_health$HeartBorn_2223)
```
Make sense of these data, turn it into factors.
```{r}
ped_health <- ped_health %>%
  mutate(heart_born = case_when(
    HeartBorn_2223 == 1 ~ "No",
    HeartBorn_2223 == 2 ~ "Ever told, NOT Born",
    HeartBorn_2223 == 3 ~ "Ever told, Born",
    HeartBorn_2223 == 99 ~ "Missing"
  )) %>%
  mutate(heart_born = factor(heart_born, levels = c("No", "Ever told, NOT Born", "Ever told, Born", "Missing")))
```
Visualize this distribution
```{r}
ggplot(ped_health, aes(x = heart_born)) +
  geom_bar() +
  labs(title = "Distribution of Cardiac Illness at Birth", x = "Cardiac Illness Status", y = "Count") +
  theme_minimal() +
  scale_x_discrete(drop = FALSE) # to ensure all factor levels are shown)
```
Making binary distribution of those born with cardiac illness
```{r}
ped_health <- ped_health %>%
  mutate(
    CHD_binary = case_when(
      HeartBorn_2223 == 3 ~ "Yes",
      HeartBorn_2223 %in% c(1, 2) ~ "No",
      HeartBorn_2223 == 99 ~ NA_character_
    ),
    CHD_binary = factor(CHD_binary, levels = c("No", "Yes"))
  )
table(ped_health$CHD_binary)
```
Making binary distribution of those with ADHD

```{r}
ped_health <- ped_health %>%
  mutate(
    adhd_binary = case_when(
      K2Q31A == 1 ~ "Yes",
      K2Q31A == 2 ~ "No",
      K2Q31A %in% c(95, 99) ~ NA_character_
    ),
    adhd_binary = factor(adhd_binary, levels = c("No", "Yes"))
  )
```

```{r}
table(ped_health$K2Q31A)
```

