---
title: "R Notebook"
output: html_document
---
### CHD and ADHD comparison with SES factors
```{r}
# for data manipulation and visualization
library(tidyverse)

# to help set the working directory and make this script reproducible across different machines
library(here)

# for the strobe diagram
library(DiagrammeR)
library(DiagrammeRsvg) 
library(glue)
library(rsvg)

library(tableone) 
library(broom)
```

```{r}
# set the working directory to the location of this script
here::here()
```

# load the data and check distribution of variables starting with poverty level
```{r}
ped_health <- read_csv(here("ped_health.csv"))
summary(ped_health$povlev4_2223)
table(ped_health$povlev4_2223)
```



#What do these numbers mean? Let's visualize.
```{r}
#1 -"0-99% of poverty level"
#2 -"100-199% of poverty level"
#3 "200-399% of poverty level"
#4 "400% or more of poverty level

ped_health<- ped_health %>%
  mutate(povlev4_2223 = factor(povlev4_2223, 
                               levels = c(1, 2, 3, 4), 
                               labels = c("0-99% of poverty level", 
                                          "100-199% of poverty level", 
                                          "200-399% of poverty level", 
                                          "400% or more of poverty level")))

ggplot(ped_health, aes(x = povlev4_2223, fill=adhd_binary))+
  labs(x="Poverty Level") +
  geom_bar()
```
```{r}
unique(ped_health$povlev4_2223)
```
#Table of people
```{r}
table(ped_health$povlev4_2223, ped_health$adhd_binary, useNA = "always")

```
#Let's see graph in proportions
```{r}
ggplot(ped_health, aes(x = povlev4_2223, fill=adhd_binary))+
  labs(x="Poverty Level") +
  geom_bar(position="fill")+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal()

prop_table <- ped_health %>%
  group_by(povlev4_2223, adhd_binary) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(povlev4_2223) %>%
  mutate(prop = n / sum(n))

print(prop_table)
```
### Wow.. while there are more children with ADHD in higher income groups when looking at raw counts, the proportion of children with ADHD is actually higher in lower income groups.

#Now let's create a contigency table for a chisquare test.
```{r}
chisq.test(prop_table$n, 
           p = prop_table$prop, 
           rescale.p = TRUE, 
           simulate.p.value = TRUE, 
           B = 10000)

```

